---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First I unzip the data using the unzip function. Then load the activity data from your current working directory using read.csv. The data will be assigned to variable activity.data.

```{r unzip, echo = TRUE, results = "hide"}
unzip("./activity.zip")
activity.data <- read.csv("./activity.csv", colClasses = c("numeric", "character", "numeric"))
```

Then I change the date to the Posixlt format for ease of use later in the analysis.

```{r updatedate, echo = TRUE, results = "hide"}
unzip("./activity.zip")
activity.data$date <- as.POSIXlt(activity.data$date)
```

## What is mean total number of steps taken per day?
In order to determine the sum of steps each day I use the tapply function and assign it to sum.day.

```{r sumday, echo = TRUE, results = "hide"}
sum.day <- tapply(activity.data$steps, as.POSIXct(activity.data$date), sum, na.rm = TRUE)
```

Now I will generate the histogram to show the distribution of total daily steps.

```{r sumhisto, echo = TRUE, fig.height = 4}
hist(sum.day, main = "Histogram - Total Daily Steps", xlab = "Total Daily Steps")
```

```{r sumdaycenters, echo = TRUE, results = "hide"}
sum.mean <- mean(sum.day)
sum.median <- median(sum.day)
```

The mean of total daily steps is: `r sum.mean`, and the median is: `r sum.median`.

## What is the average daily activity pattern?
I will now generate the average daily steps for each 5 minute interval.

```{r intervalaverage, echo = TRUE, results = "hide"}
sum.interval <- tapply(activity.data$steps, activity.data$interval, mean, na.rm = TRUE)
```

This is a time series of the results from the above computation.

```{r timeseries, echo = TRUE, fig.height = 4}
plot(sum.interval, type = "l", xlab = "Interval", ylab = "Average Daily Steps", main = "Average Daily Steps Time Series")
```

```{r intervalaveragemax, echo = TRUE, results = "hide"}
max.interval <- max(sum.interval)
interval.max <- as.numeric(names(sum.interval[sum.interval = max.interval]))
```
There is an interesting peak in the time series. Let's see what the value is, and for which interval it is.  

The value is `r max.interval` for interval `r interval.max`.

## Imputing missing values
I saw some missing values in the dataset. All of the previous calculations have removed the NAs. That might distort the analysis. Now I will update the NAs to allow for a more robust analysis.  

```{r missingcount, echo = TRUE, results = "hide"}
missing.count <- sum(is.na(activity.data))
```

First, I calculate the number of missing values to be `r missing.count`. I impute a value for each them, that is a fancy way of saying I will give them a value. This is done by way of random sampling from the non-missing values. Note I am imputing the data into another variable, so that I can have my original dataset preserved for reference.

```{r imputemissing, echo = TRUE, results = "hide"}
activity.impute <- activity.data
activity.impute$steps[is.na(activity.impute$steps)] <- sample(activity.impute$steps[!is.na(activity.impute$steps)], missing.count, replace = TRUE)
```

I revisit an earlier analysis to compare the impute data set to the original. The histogram of total daily steps will do nicely for comparison. Using the same steps from above I arrive at the below 2 histograms for imputed and not data.

```{r imputesumday, echo = TRUE, results = "hide"}
sum.day.impute <- tapply(activity.impute$steps, as.POSIXct(activity.impute$date), sum)
```

```{r sumhistoimpute, echo = TRUE, fig.height = 4, fig.width = 10}
par(mfrow = c(1,2), mar = c(4,4,4,4))
hist(sum.day.impute, main = "Histogram - Total Daily Steps (Imputed)", xlab = "Total Daily Steps")

hist(sum.day, main = "Histogram - Total Daily Steps", xlab = "Total Daily Steps")
```

```{r sumdaycentersimpute, echo = TRUE, results = "hide"}
sum.mean.impute <- mean(sum.day.impute)
sum.median.impute <- median(sum.day.impute)
mean.difference <- abs(sum.mean - sum.mean.impute)
median.difference <- abs(sum.median - sum.median.impute)
difference.text <- if(mean.difference > 0 || median.difference > 0) { "There is a difference in center between the original and imputed datasets. Imputing normalizes the dataset."} else {"WOW! There is no difference in the center between the original and imputed datasets. Imputing doesn't change anything."}
```

The means are `r mean.difference` steps apart, and the medians are `r median.difference` steps apart. `r difference.text`.


## Are there differences in activity patterns between weekdays and weekends?
